---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SimpleLogistic
[![Codecov test coverage](https://codecov.io/gh/yanax726/Simple_Logistic/graph/badge.svg)](https://app.codecov.io/gh/yanax726/Simple_Logistic)

## Introduction

**SimpleLogistic** is an R package I developed to simplify the process of performing logistic regression analysis, especially tailored for health data science applications. It provides easy-to-use functions for fitting logistic regression models, generating diagnostic plots, and visualizing predicted probabilities. Whether you're a student just getting started with logistic regression or a practitioner analyzing health outcomes, this package aims to make your workflow more efficient and intuitive.

## Features

- **Easy Model Fitting**: Quickly fit logistic regression models using a simple function.
- **Diagnostic Plots**: Generate basic diagnostic plots to check your model.
- **Prediction Visualization**: Plot predicted probabilities against actual outcomes.
- **Interaction Terms**: Easily include interaction terms in your models.
- **Performance Benchmarking**: Compare the efficiency of simple_logistic() with R's base glm() function using the bench package.
- **Comprehensive Vignette**: Access a detailed guide with examples and explanations to help you make the most of the package.
- **Compatibility with Base R**: Works smoothly with base R functions for further analysis.

## Installation

You can install **SimpleLogistic** from GitHub:

```{r, eval = FALSE}
# Install devtools if you haven't already
install.packages("devtools")

# Load devtools
library(devtools)

# Install SimpleLogistic from GitHub
install_github("yourusername/SimpleLogistic")
```
Make sure to replace "yourusername/SimpleLogistic" with the actual path to your GitHub repository.

## Getting Started
### Load the package and the example dataset:
``` {r}
library(SimpleLogistic)

# Load the example dataset
data("health_data")

# Look at the first few rows
head(health_data)
```
#### The health_data dataset includes:
**outcome**: Binary outcome variable (0 or 1)
**age**: Age in years
**bmi**: Body Mass Index
**treatment**: Treatment group ("A" or "B")

## Usage
### Fitting a Logistic Regression Model
Use the simple_logistic() function to fit a model:
```{r}
# Fit the model
model <- simple_logistic(outcome ~ age + bmi + treatment, data = health_data)
# View the summary
summary(model$fit)
```
This function simplifies the model fitting process by wrapping around the glm() function, allowing you to focus more on interpreting results rather than setting up the model.

### Generating Diagnostic Plots
After fitting your model, it's essential to check how well it fits the data. The diagnostic_plots() function helps you create two key diagnostic plots.
```{r}
# Generate diagnostic plots
plots <- diagnostic_plots(model)

# Residuals vs Fitted Values
print(plots$residuals_vs_fitted)

# ROC Curve
print(plots$roc_curve)
```
These plots help you assess the goodness-of-fit and the model's ability to discriminate between outcome classes.

### Visualizing Predicted Probabilities
Understanding the distribution of predicted probabilities can provide valuable insights into your model's performance. Use the plot_predictions() function to visualize this.
``` {r}
# Plot predicted probabilities
plot_predictions(model)
```
This histogram shows how predicted probabilities are distributed across actual outcome classes, helping you gauge the model's effectiveness.

### Efficiency Benchmarking
Efficiency matters, especially with large datasets. Here's how you can benchmark the performance of simple_logistic() against R's base glm() function using the bench package.
```{r}
# Create a larger dataset for benchmarking
set.seed(123)
n_large <- 10000
age_large <- rnorm(n_large, mean = 50, sd = 10)
bmi_large <- rnorm(n_large, mean = 25, sd = 4)
treatment_large <- sample(c("A", "B"), n_large, replace = TRUE)
logit_large <- -5 + 0.05 * age_large + 0.1 * bmi_large + ifelse(treatment_large == "A", 0.5, -0.5)
prob_large <- 1 / (1 + exp(-logit_large))
outcome_large <- rbinom(n_large, size = 1, prob = prob_large)
health_data_large <- data.frame(
  outcome = outcome_large,
  age = age_large,
  bmi = bmi_large,
  treatment = treatment_large
)

# Benchmarking the simple_logistic function against glm
bench_simple_logistic <- bench::mark(
  simple_logistic_model = simple_logistic(outcome ~ age + bmi + treatment, data = health_data_large),
  glm_model = glm(outcome ~ age + bmi + treatment, data = health_data_large, family = binomial()),
  iterations = 10,
  check = FALSE  # Disable result comparison
)

print(bench_simple_logistic)
```
Note: We set check = FALSE because simple_logistic() and glm() return different types of objects. This setting allows us to focus solely on benchmarking the performance without worrying about result equivalence.

#### Interpreting the Results:
The benchmark will provide metrics like median execution time, minimum and maximum times, memory allocations, and garbage collections. You can compare how simple_logistic() stacks up against glm() in terms of speed and memory usage.

## Advanced Usage
### Including Interaction Terms
Sometimes, the relationship between a predictor and the outcome might depend on another predictor. You can include interaction terms in your model formula to explore these dynamics.
``` {r}
# Fit a model with an interaction term
model_interaction <- simple_logistic(outcome ~ age * treatment + bmi, data = health_data)

# Summarize the model
summary(model_interaction$fit)
```
In this example, age * treatment includes both the main effects and their interaction. This allows you to see if the effect of age on outcome changes depending on the treatment group.

### Making Predictions on New Data
After fitting a model, you might want to make predictions on new observations. Here's how you can do that.
``` {r}
# New data for prediction
new_data <- data.frame(
  age = c(35, 45, 55),
  bmi = c(24, 28, 31),
  treatment = c("A", "B", "A")
)

# Predict probabilities
new_data$predicted_prob <- predict(model$fit, newdata = new_data, type = "response")

# View predictions
print(new_data)
```
This adds a predicted_prob column to your new_data dataframe, showing the estimated probability of outcome being 1 for each new observation.

### Comparing with Base R Functions
It's always a good practice to ensure that your custom functions are producing results consistent with established base R functions. Here's how you can compare the coefficients from simple_logistic() with those from glm() to verify their consistency.
``` {r}
# Fit a model using glm()
model_glm <- glm(outcome ~ age + bmi + treatment, data = health_data, family = binomial())

# Compare coefficients
all.equal(coef(model$fit), coef(model_glm))
```
The all.equal() function checks if the coefficients from both models are nearly identical, ensuring that simple_logistic() is functioning correctly.

## Dataset
The package includes a simulated dataset health_data with:

- **outcome:** Binary outcome variable (0 or 1)
- **age:** Age in years
- **bmi:** Body Mass Index
- **treatment:** Treatment group ("A" or "B")
You can load it using:
```{r}
data("health_data")
head(health_data)
```
## Dependencies

SimpleLogistic uses the following R packages:
- **ggplot2** for plotting.
- **pROC** for ROC curve analysis.
- **rlang** For tidy evaluation.
Additionally, for benchmarking and documentation:

- **bench**  For performance benchmarking.
- **knitr and rmarkdown** For creating vignettes and documentation.
- **testthat** For running tests.
You can install them with:
```{r, eval = FALSE}
install.packages(c("ggplot2", "pROC", "rlang", "bench", "knitr", "rmarkdown", "testthat"))

```

## Contributing
If you have suggestions for improvements, find any issues, or want to contribute to the development of SimpleLogistic, feel free to open an issue or submit a pull request on GitHub.

## License
This project is licensed under the MIT License.

I hope SimpleLogistic is helpful for your logistic regression analyses, especially if you're just getting started with R and statistical modeling.

For more examples and detailed usage, check out the package vignette:
```{r, eval = FALSE}
browseVignettes("SimpleLogistic")
```

